---
title: "Introduction to Plotting and Model Creation"
author: "Yogindra Raghav"
date: "September 20, 2018"
output: pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("plotrix")
dat<-read.table("C:/Users/Yogindra Raghav/Downloads/mtmv_data_10_12.csv", header = TRUE,
                sep = ",")
attach(dat)
```

This chunk of code reads the data in by allowing users to choose csv file and then uses attach() function for easier access of column headers. 

```{r, eval = TRUE}
model_namebi<-lm(vote_share~mshare)
m10<-lm(vote_share~mshare)
a10<-m10$coefficients[1]
b10<-m10$coefficients[2]

m12<-lm(vote_share_12~mshare_12)
a12<-m12$coefficients[1]
b12<-m12$coefficients[2]
plot(mshare,vote_share, pch=20, cex=1.0, ylim=c(0, 100), xlab="Tweet Share", 
     ylab="Vote Share")

abline(a10, b10, lwd=2, col=2)

``` 

This code runs the simple linear regression model for a given data set using the lm() function which stands for linear model. 

These lines are responsible for first making the scatter plot using the plot() function and then adding a line to a plot with the abline() function. 



ORIGINIAL FROM MDSR BOOK: 

```{r results = "asis", warning = FALSE}
library(xtable) 
library(mdsr) 
options(xtable.comment= FALSE) 
mod = lm(cesd ~ mcs + sex, data = HELPmiss) 
xtable(mod) 
```


MY OWN ANALYSIS USING IRIS DATASET: 

```{r results = "asis", warning = FALSE}

options(xtable.comment= FALSE)
mod = lm(Sepal.Length~Petal.Length, data = iris)
xtable(mod)
```



